---
title: "HW1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
data <- read_csv("HW1 Data.csv")
```

```{r SHARK}
#rename columns
names(data)<-c("pixel", "cost", "tuna", "shark")

#calculate alpha
shark.a<-1/(sum(data$shark))^0.6 #Shark Alpha 0.01777361
tuna.a<-1/(sum(data$tuna))^0.6 #Tuna Alpha 0.01604304

```
a)
b)Max for sharks
```{r}
max.shark<-data %>%
  mutate(cost.per.suit= cost/shark) %>% 
  arrange(cost.per.suit) %>% 
  mutate(pixel.cost = cumsum(cost)) %>%
  mutate(frac = ifelse(pixel.cost<=200, 1, (200-lag(pixel.cost))/cost)) %>% 
  mutate(total.cost = (cost)*frac) %>%
  mutate(total.cost = cumsum(total.cost)) %>% 
  filter(total.cost<=200 & frac>=0) %>% 
  mutate(shark.suit = shark*frac) %>% 
  mutate(tuna.suit = tuna*frac) 
  

##Calculate the probability of survival 
shark.a*((sum(max.shark$shark.suit))^0.6) #0.5970874
tuna.a*((sum(max.shark$tuna.suit))^0.6) #0.5956662

write.csv(max.shark, "Shark Priority.csv", row.names = FALSE)

```
c)Max for Tuna 

```{r}
max.tuna<-data %>%
  mutate(cost.per.suit= cost/tuna) %>% 
  arrange(cost.per.suit) %>% 
  mutate(pixel.cost = cumsum(cost)) %>%
  mutate(frac = ifelse(pixel.cost<=200, 1, (200-lag(pixel.cost))/cost)) %>% 
  mutate(total.cost = (cost)*frac) %>%
  mutate(total.cost = cumsum(total.cost)) %>% 
  filter(total.cost<=200 & frac>=0) %>% 
  mutate(shark.suit = shark*frac) %>% 
  mutate(tuna.suit = tuna*frac) 
  

##Calculate the probability of survival 
shark.a*((sum(max.tuna$shark.suit))^0.6) #0.5884504
tuna.a*((sum(max.tuna$tuna.suit))^0.6) #0.6075198


```

d)Random selection of pixels within the budget. Repeat 10 times 
```{r}
random<-data[sample(nrow(data)),] %>% 
  mutate(pixel.cost = cumsum(cost)) %>%
  mutate(frac = ifelse(pixel.cost<=200, 1, (200-lag(pixel.cost))/cost)) %>% 
  mutate(total.cost = (cost)*frac) %>%
  mutate(total.cost = cumsum(total.cost)) %>% 
  filter(total.cost<=200 & frac>=0) %>% 
  mutate(tuna=tuna*frac) %>% 
  mutate(shark=shark*frac)

tuna.a*((sum(random$tuna))^0.6)
shark.a*((sum(random$shark))^0.6)
```

e)Weighted 
(Shark suit/cost * W)+(Tuna Suit/cost * (1-W))

Not knowing which species is more important to society, suppose you place weight w on survival of
sharks and weight (1-w) on survival of tuna. By varying the weights, w, trace out the efficient
frontier of attainable survival probabilities for both species. 
```{r}
w<-0.2
weight<-data %>% 
  mutate(weight.suit= (shark/cost)*w+(tuna/cost)*(1-w)) %>%
  arrange(-weight.suit) %>%
  mutate(pixel.cost = cumsum(cost)) %>%
  mutate(frac = ifelse(pixel.cost<=200, 1, (200-lag(pixel.cost))/cost)) %>% 
  mutate(total.cost = (cost)*frac) %>%
  mutate(total.cost = cumsum(total.cost)) %>% 
  filter(total.cost<=200 & frac>=0) %>% 
  mutate(tuna=tuna*frac) %>% 
  mutate(shark=shark*frac)

tuna.a*((sum(weight$tuna))^0.6)
shark.a*((sum(weight$shark))^0.6)
```

